---
title: "Class 11: Structural Bioinformatics Pt. 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The PDB Database

The [PDB](http://www.rcsb.org/) is a large repository for detailed data on various proteins.

Let's explore the PDB by examining a .csv download with detailed statistics on the PDB.

>Q1: Download a CSV file from the PDB site (accessible from “Analyze” -> “PDB Statistics” >
“by Experimental Method and Molecular Type”. Move this CSV file into your RStudio project
and determine the percentage of structures solved by X-Ray and Electron Microscopy. From
the website what proportion of structures are protein? Aim to have a rendered GitHub document
with working code that yields your answers.

```{r}
db <- read.csv("PDB Summary Data.csv")

#Percent that are X-Ray & Electron Microscopy, etc.
(db$Total / sum(db$Total)) * 100

#Percent that are proteins
(sum(db$Proteins) / sum(db$Total)) * 100
```

We just installed data pasta by runing install.packages("datapasta"). This is a tool that helps copy and paste raw data into R without having to do additional formatting. **Advantage:** This was much faster than downloading the data and importing it into our R project as we did before.

```{r}
library(datapasta)


data.frame(stringsAsFactors=FALSE,
   Experimental.Method = c("X Ray", "NMR", "Electron Microscopy", "Other",
                           "Multi Method", "Total"),
              Proteins = c(126880, 11062, 2277, 256, 129, 140604),
         Nucleic.Acids = c(2012, 1279, 31, 4, 5, 3331),
    Protein_NA.Complex = c(6547, 259, 800, 6, 2, 7614),
                 Other = c(8, 8, 0, 13, 1, 30),
                 Total = c(135447, 12608, 3108, 279, 137, 151579)
)


```

>Q2: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?

As of 5/1/19 - there are 1,157 proteins. Results are visible (here)[http://www.rcsb.org/pdb/results/results.do?tabtoshow=Current&qrid=727B9D8A]


##Section 2: Visualizing the HIV-1 protease protein in VMD viewer.
The next instruction, is to download the .pdf file for HIV-1 proteae with identifier 1HSG and import it into this project. 

In order to do this, we head to install the VMD application available here(https://www.ks.uiuc.edu/Development/Download/download.cgi)


##Section 3: Utilizing RStudio and Bio3D to do the same thing within R environment & more! Dr. Grant has provided us two overarching directions on this project. 1. Extract *Protein Only* PDB, and 2. Extract *drug/ligand* only PDB

First, we will load the bio3d package
```{r}
library(bio3d)
```

We will now pull the data from our pdb file using **read.pdb()** and assign it to the variable pdb. Note how we can just do this right away without having to download the package. We'll also print it - just as a cross-check.
```{r}
pdb <- read.pdb("1hsg")

pdb
```

>Q6. How many amino acid residues are there in this pdb object and what are the two nonprotein residues?
There are 198 total AA residues. The two non-protein resiudes are HOH (water) and MK1.

Note that the attributes (+ attr:) of this object are listed on the last couple of lines. To find the
attributes of any such object you can use:
```{r}
attributes(pdb)
```


o access these individual attributes we use the dollar-attribute name convention that is
common with R list objects. For example, to access the atom attribute or component use 
```{r}
head(pdb$atom)
```

> Q7. What type of R object is pdb$atom? HINT: You can always use the str() function to get a
useful summery of any R object. Per below, we can see that it's a data frame. 

```{r}
pdb$atom

str(pdb$atom)
```


>Q8. Use the Bio3D write.pdb() function to write out a protein only PDB file for viewing in
VMD. Also write out a second separate PDB file for the ligand with residue name MK1
Atom selection is done via the function **atom.select()**. NOte in the example below how there is exactly 3x as many XYZ indices as atoms. This is because each atom should have it's own X, Y, and Z coordinate to determine its 3d arrangement in space.

```{r}
inds <- atom.select(pdb, "protein")
```

```{r}
pdb$atom[inds$atom,]
```

Notice how there's nothing now but protein - no water, etc.
```{r}
prot.pdb <- atom.select(pdb, "protein", value=TRUE)

write.pdb(prot.pdb, file="1hsg_protein.pdb")
```

We will now do the same thing for ligand
```{r}
ligand.pdb <- atom.select(pdb, "ligand", value=TRUE)

write.pdb(ligand.pdb, file="1hsg_ligand.pdb")
```

We just opened both of our new .pdb files in **VMD** to verify they looked right. Note how we can create new files to read elsewhere. In this case, we wanted files that just showed protein & just showed ligands.

##Section 4 - Working with multiple PDB files

First step is to install MUSCLE and moving the .exe file directly into this project.

Next, Dr. Grant wants us to try to align multiple structures. 
```{r}
# Download some example PDB files
ids <- c("1TND_B","1AGR_A","1TAG_A","1GG2_A","1KJY_A","4G5Q_A")

files <- get.pdb(ids, split = TRUE)
```

##Section 5 - 

Step 1 - Search and retrieve 
